### 2014-03-01 paper/related.tex
First pass at related work section.
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

@@

Gotta talk about secure multi-execution

>>>>

Our paper follows a long tradition of papers on information flow
control.  Our information flow control system is mostly closely related
to the coarse-grained information systems used in operating systems such
as Asbestos~\cite{efstathopoulos:asbestos} and
HiStar~\cite{Zeldovich:2006}, as well as language-level
\emph{floating-label IFC systems} such as LIO~\cite{lio},
Breeze~\cite{Hritcu:2013:YIB:2497621.2498098}, where there is a label
associated with threads of execution which increases monotonically.
Our treatment of termination-sensitive and termination-insensitive interference
originates from Smith and Volpano~\cite{Smith:Volpano:MultiThreaded,Volpano:1997:ECF:794197.795081}.
While this work often considers the interaction of information flow
control with respect to various language features (e.g.,
exceptions~\cite{Hritcu:2013:YIB:2497621.2498098}), none of these
systems attempt to be generic in the choice of target language.

One of the primary motivations behind this paper is constructing a
practical information flow control system for JavaScript.  Some of
these systems attempt to perform fine-grained information flow of
JavaScript, e.g. Hedin and Sabelfield's work~\cite{Hedin:2012} on JSFlow~\cite{JSFlow}
and \Red{some other stuff here}.  While fine-grained information flow
control can result in very precise analysis of legacy code, it comes
at the cost of high theoretical complexity (the system must accomodate
the entirety of JavaScript's semantics) and runtime performance cost.
There are attempts to overcome this performance barrier by various means:
Austin and Flanagan~\cite{Austin:Flanagan:PLAS09} handle this problem
by using a \emph{sparse information labeling} scheme which reduces the
cost of managing labels when they are not needed, while FlowFox~\cite{DeGroef:2012}
uses an entirely different technique of \emph{secure multi-execution} to
enforce non-interference. \Red{Probably more stuff here.}

The constructs in our IFC language, as well as the behavior of
inter-thread communication, is reminiscent of distributed systems
like Erlang.~\cite{Armstrong03makingreliable}  There is a close
correspondence: in distributed systems, isolation is required due to
physical constraints; in information flow control, isolation is
required to enforce noninterference.  Papagiannis et al.~\cite{Papagiannis_enforcinguser}
built an information flow control system on top of Erlang which is quite
similar to ours.  However, they do not use a floating label system (processes
can find out when sending a message failed due to a forbidden information flow),
and they include no security proofs.

There appears to be limited literature in general techniques for
retrofitting information flow control to arbitrary languages; however,
one time-honored technique is defining a fundamental calculus, for which
other languages can be desugared into.  Abadi et al.~\cite{abadi+:core}
motivate their core calculus of dependency by showing how various
previous systems can be encoded in it; Tse and
Zdancewic~\cite{Tse:Zdancewic:ICFP04}, in turn, show how this calculus
can be encoded in System F via parametricity.  However, this work is
primarily focused on static information flow control systems.

Our definition of isomorphism of information flow control languages in
Section~\ref{sec:concrete} is closely related to the idea of
\emph{bisimulation}~\cite{Milner:1989:CC:534666} and process calculi,
although as our system is deterministic, we do not use the theory in any
deep way.

It has long been observed that information flow control and monads are
closely related.  This observation has been used by Abadi et
al.~\cite{abadi+:core} to structure access to private data by protecting
it using a lattice of monads, one per security level.  Our system uses
monads in a conceptually different manner, where there is a single monad
representing the information flow control computational effects; this
is inline with the use of monads in other systems.~\cite{Harrison05,lio,Devriese:2011}.
Devriese and Piessens~\cite{Devriese:2011} utilized monad
transformers as an essential component in their work.  However, their use
of monad transformers is to explain when extra constraints on a non-information
flow control aware base monad should be applied (i.e., during lifting).
Harrison and Hook~\cite{Harrison05} define a notion of \emph{atomic noninterference},
which is preserved through monad transformer application.  However, it is not
clear what the relationship between this notion and traditional noninterference is,
and in their paper they prove only a weaker property of no write down.

\cut{
\begin{itemize}
    \item Monads and IFC (Abadi~\cite{Abadi+:Core}, Tse~\cite{Tse:Zdancewic:ICFP04}, Harrison-Hook~\cite{Harrison05}, Crary~\cite{Crary:2005}, Devriese-Piessens~\cite{Devriese:2011})
    \item \Red{Operating systems approaches to IFC, which use coarse grained (Asbestos~\cite{efstathopoulos:asbestos}, HiStar~\cite{Zeldovich:2006})}
    \item \Red{Floating-label IFC systems, look at LIO paper (LIO~\cite{lio}, Breeze~\cite{Hritcu:2013:YIB:2497621.2498098})}
    \item JavaScript IFC related work, look at BrowBound. (ADSafe, xBook \Red{needs lookup})  Classify approaches into fine-grained (Hedin-Sabelfield~\cite{Hedin:2012}+JSFlow~\cite{JSFlow}, FlowFox~\cite{DeGroef:2012}, ConDOM~\cite{ConDOM}, Austin-Flanagan (sparse information labeling)~\cite{Austin:Flanagan:PLAS09}) or coarse-grained (BFlow~\cite{Yip:2009:PBS}, see Deian)
\item Secure multi-execution (FlowFox~\cite{DeGroef:2012}, Russo~\cite{Jaskelioff:SME})
\item Distributed programming languages (Cloud Haskell) plus IFC (Erlang (they check IFC when sending messages, no security proofs, assumes true isolation, not a floating labels)~\cite{Papagiannis_enforcinguser})
\end{itemize}

@@

Gotta talk about secure multi-execution
}
